### 2016-04-13 text/ms.Rmd
Many edits to the text

Population status and trend are two of the most fundamental values to quantify
in fisheries management [e.g. TODO NEW REFS]. However, fisheries managers are
often faced with reconciling multiple uncertain and potentially conflicting
estimates of status and trend [e.g. @brodziak2010; @branch2011; @deroba2015].
For example, some models may suggest a population is at risk and declining in
abundance while others may suggest it is not at risk and stable.

One solution is to take the average or weighted average of several model
predictions, i.e. an ensemble. Such ensembles are typically more accurate and
less biased than individual model estimates and can integrate uncertainty in
model structure, initial conditions, and parameter estimation
[@dietterich2000; @araujo2007]. Ensembles are superior to individual models
for at least three reasons: statistically by averaging across models and
therefore being less likely to pick the "wrong" model, computationally by
reducing the risk of getting stuck in a local optima, and representationally
by expanding the range of hypotheses explored [@dietterich2000]. This approach
forms the basis of many machine learning methods [e.g. @dietterich2000], has
helped reconcile climate forecasts from dozens of models [e.g. @murphy2004;
@tebaldi2007; @ipcc2013], and even improved early warning signs of malaria
outbreaks [@thomson2006]. In ecology, ensemble methods are sometimes used to
improve species distribution modelling [e.g. @araujo2007; @breiner2015] and
indeed have been used to combine estimates of population status and trend
[e.g. @brodziak2010].

@@

Therefore, a number of models have been proposed to assess $B/B_\mathrm{MSY}$
based on the limited data available for the majority of fish stocks: (1)
a time series of the total biomass of catch and (2) a basic understanding of
population productivity [e.g. @vasconcellos2005; @martell2013]. Recently,
@rosenberg2014 investigated the performance of four data-limited models
through a large-scale simulation experiment. Three of these models are based
on Schaefer (logistic) biomass dynamics and one is an empirical model fitted
to more data-rich stock assessment output. The four models frequently
disagreed about population status, no one model had strong performance across
all fish stocks, and some models performed better than others depending on
circumstances (e.g. Fig. \ref{motivate}).

@@

<!-- TODO Add a summary of the methods here  -->

@@

\noindent where $\hat{B}_{t+1}$ represents predicted biomass at time $t$ plus
one year, $B_t$ represents biomass at time $t$, $r$ represents the intrinsic
population growth rate, $B_0$ represents unfished biomass or carrying capacity
$K$, and $C$ represents catch. The fourth model is an empirically derived
model based on the RAM Legacy Stock Assessment Database. @rosenberg2014
provide a full background on these four methods and code to fit all the models
is available in an accompanying package `datalimited` for the statistical
software R TODO ADD URL. As described in the Introduction, the performance of
these models at estimating $B/B_\mathrm{MSY}$ is variable. In summary:

@@

## Simulated data set to build the ensemble

@@

## Superensemble models

@@

We used the mean or slope of $B/B_\mathrm{MSY}$ as the response variable and
the predictions from the individual models <!--, and the spectral values, fix
this explanation --> as predictors in our superensemble models. When modelling
mean $B/B_\mathrm{MSY}$ --- a ratio bounded at zero --- we fit the
superensemble models in log space and exponentiated the predictions. For the
estimates of $B/B_\mathrm{MSY}$ slope, which are not bounded at zero, we fit
superensemble models on the natural untransformed scale.

We compared an ensemble average and three superensembles of varying
complexity: a linear model with two-way interactions, a random forest, and
a boosted regression tree. We describe these models as estimating
$\hat{\theta}$, which represents either the ensemble estimated $\log
B/B_\mathrm{MSY}$ or slope of $B/B_\mathrm{MSY}$. The individual model
estimates of $\log B/B_\mathrm{MSY}$ or slope of $B/B_\mathrm{MSY}$ are
represented as $b$ for models $i$ $1$ through $n$. The subscript $j$ indexes
the fisheries from $1$ through $m$. The ensemble average was calculated as:

@@

\begin{equation}
\hat{\theta_j} = \beta_0 + \beta_1 \hat{b}_{j,0} + ... +
\beta_{1,2} \hat{b}_{j,1} \hat{b}_{j,2} + ... +
\epsilon_j,
\quad
\mathrm{Normal}(0, \sigma^2)
\quad \mathrm{for}\ j  =  1,...,m
\end{equation}

Our two machine learning superensemble models, a random forest and a boosted
regression tree, were based on regression trees. Regression trees sequentially
determine what value of a predictor best splits the response data into two
'branches' based on a loss function [@breiman1984]. In random forests,
a series of regression trees are built on a random subset of the data and
random subset of the covariates of the model [@breiman2001]. In generalized
boosted models (GBMs), each subsequent model is fit to the residuals from the
previous model; data points that are fit poorly in a given model are given
more weight in the next model [@elith2008]. Random forests and GBMs can
provide strong predictive performance and fit highly non-linear relationships
[@elith2008; @hastie2009]. We fit random forest models with the `randomForest`
package [@liaw2002] for R with the default argument values. We fit boosted
regression tree models with the `gbm` package [@ridgeway2015] for R. Based on
initial tuning with the **caret** R package [@caret2008], we chose to fit GBMs
with 2000 trees, an interaction depth of $6$, a learning rate (shrinkage
parameter) of $0.01$, and all other arguments at their default values:

TODO ADD EQN

We also tested our ensemble methods on the RAM Legacy Stock Assessment
Database [@ricard2012] --- a compilation of stock-assessment output from
hundreds of exploited marine populations around the world. Our analysis of the
stock-assessment database was based on version 2.5. After removing stocks for
which at least one of the individual models described below did not converge
(121), this database included `r ram_stocks_n` stocks.

@@

<!-- Superensemble models allow us to incorporate additional covariates and
potentially leverage interactions between these covariates and individual model
predictions. Additional covariates could be, for example, life-history
characteristics, information on exploitation patterns, or statistical properties
of the data. For simplicity, and since many possible covariates, such as fishing
effort dynamics, would not be known for many real-world data-limited scenarios,
we added only one set of additional covariates: spectral properties of the catch
time series. Spectral analysis decomposes a time series into the frequency
domain and provides a means of describing the cyclical shape of the catch series
that is independent of time series length (except in affecting precision) and
absolute magnitude of catch. We fit spectral models to the scaled catch time
series (catch divided by maximum catch) with the `spec.ar` function in R and
recorded representative short- and long-term spectral densities at frequencies
of 0.20 and 0.05, which correspond to 5- and 20-year cycles. -->

<!--*TODO add new supp. fig.?*-->

## Testing model performance

A critical component to predictive modelling is to evaluate the performance of
a model on new data [@hastie2009]. We used repeated three-fold cross validation
to test predictive performance: we randomly divided the dataset into three sets,
built superensemble models on two-thirds of the data, and evaluated predictive
performance on the remaining third. We repeated this across each of the three
splits and then repeated the whole procedure 50 times to account for bias that
may result from any one set of validation splits. In the simulated dataset,
there were 10 replicates of each unique combination of simulation parameters
that differed only in stochastic variability. Since the dynamics of these
populations were often similar, we grouped these stocks in the cross-validation
process into either the training or testing split. We also validated the models
using the RAM Legacy Stock Assessment database. In this case, we used
superensembles trained on the entire simulation dataset. However, since mPRM is
built on the same stock-assessment database, we applied three-fold
cross-validation to the RAM Legacy Stock Assessment data underlying the mPRM
model so that the dataset that mPRM was trained with (for the individual model
and superensemble) was separate from the dataset it was tested on.

Predictive performance can be evaluated with metrics that represent a variety of
modelling goals. For continuous response variables such as the mean and slope of
population status, performance metrics often measure some form of bias,
precision, accuracy (a combination of bias and precision), or the ability to
correctly rank or correlate across populations [e.g. @walther2005]. Here, we
measure proportional error, defined as $(\hat{b} - b)/\abs{b}$, where $\hat{b}$
and $b$ represent estimated and 'true' (or stock-assessed) mean or slope of
$B/B_\mathrm{MSY}$. We calculated median proportional error to measure bias,
median absolute proportional error to measure accuracy, and Spearman's
rank-order correlation between predicted and 'true' values to measure the
ability to correctly rank populations. When cross-validating against the RAM
Legacy Stock Assessment database, we treated the estimates from these data-rich
stock assessments as known without error. Thus, any error in the
stock-assessment estimates of the mean or slope of $B/B_\mathrm{MSY}$ also
contributes to our estimates of prediction error for each of the four
data-limited models and the ensembles. Code to reproduce our analysis and an
R package to fit the data-limited assessment models are available at
<https://github.com/datalimited/ensembles> (*Private while in review*).

@@

<!--TODO trash this?-->

@@

**Acknowledgements**: We thank members of Phase I of the working group
'Developing new approaches to global stock status assessment and fishery
production potential of the seas' who contributed to developing the data-limited
methods and simulations used in our analysis. We thank E. Jardim, F. Scott, and
J.A. Hutchings for helpful comments during the development of this project, and
R.D. Methot for comments on an earlier version of the manuscript. We thank the
Gordon and Betty Moore Foundation for funding the working group 'Applying
data-limited stock status models and developing management guidance for
unassessed fish stocks'.

@@

>>>>

Population status and trend are two of the most fundamental values to quantify
in fisheries management. However, fisheries managers are often faced with
reconciling multiple uncertain and potentially conflicting estimates of status
and trend [e.g. @brodziak2010; @branch2011; @deroba2015]. For example, some
models may suggest a population is at risk and declining in abundance while
others may suggest it is not at risk and stable.

One solution is to take the average or weighted average of several model
predictions, i.e. an ensemble. Such ensembles are typically more accurate and
less biased than individual model estimates and can integrate uncertainty in
model structure, initial conditions, and parameter estimation
[@dietterich2000; @araujo2007]. Ensembles are superior to individual models
for at least three reasons: (1) statistically by averaging across models and
therefore being less likely to pick the "wrong" model, (2) computationally by
reducing the risk of getting stuck in a local optima, and (3)
representationally by expanding the range of hypotheses explored
[@dietterich2000]. This approach forms the basis of many machine learning
methods [e.g. @dietterich2000], has helped reconcile climate forecasts from
dozens of models [e.g. @murphy2004; @tebaldi2007; @ipcc2013], and even
improved early warning signs of malaria outbreaks [@thomson2006]. In ecology,
ensemble methods are sometimes used to improve species distribution modelling
[e.g. @araujo2007; @breiner2015] and indeed have been used to combine
estimates of population status and trend [e.g. @brodziak2010].

@@

Therefore, a number of models have been proposed to assess $B/B_\mathrm{MSY}$
based on the limited data available for the majority of fish stocks: (1)
a time series of the total biomass of catch and (2) a basic understanding of
population productivity [e.g. @vasconcellos2005; @martell2013]. Recently,
@rosenberg2014 investigated the performance of four data-limited models
through a large-scale simulation experiment. Three of these models were based
on Schaefer (logistic) biomass dynamics and one was an empirical model fitted
to more data-rich stock assessment output. The four models frequently
disagreed about population status, no one model had strong performance across
all fish stocks, and some models performed better than others depending on
circumstances (e.g. Fig. \ref{motivate}).

@@

@@

\noindent where $\hat{B}_{t+1}$ represents predicted biomass at time $t$ plus
one year, $B_t$ represents biomass at time $t$, $r$ represents intrinsic
population growth rate, $B_0$ represents unfished biomass or carrying capacity
$K$, and $C$ represents catch. The fourth model is an empirically derived model
based on the RAM Legacy Stock Assessment Database [@ricard2012]. @rosenberg2014
provide a full background on these four methods and code to fit all the models
is available in an accompanying package `datalimited` for the statistical
software R <https://github.com/datalimited/datalimited> (*Private while in
review*). In summary:

@@

## Simulated dataset to build the superensemble

@@

## Building the superensemble models

@@

We used the mean or slope of $B/B_\mathrm{MSY}$ as the response variable and
the predictions from the individual models as predictors in our superensemble
models. When modelling mean $B/B_\mathrm{MSY}$ --- a ratio bounded at zero ---
we fit the superensemble models in log space and exponentiated the
predictions. For the estimates of $B/B_\mathrm{MSY}$ slope, which are not
bounded at zero, we fit superensemble models on the natural untransformed
scale.

We compared an ensemble average and three superensembles of varying
complexity: a linear model with two-way interactions, a random forest, and a
boosted regression tree. We describe these models as estimating
$\hat{\theta}$, which represents either the ensemble estimated $\log
B/B_\mathrm{MSY}$ or slope of $B/B_\mathrm{MSY}$. The individual model
estimates of $\log B/B_\mathrm{MSY}$ or slope of $B/B_\mathrm{MSY}$ are
represented as $\hat{b}$ for models $i$ $1$ through $n$. The subscript $j$
indexes the simulated fisheries from $1$ through $m$. The ensemble average was
calculated as:

@@

\begin{equation}
\hat{\theta_j} = \beta_0 + \beta_1 \hat{b}_{j,0} + ... +
\beta_{1,2} \hat{b}_{j,1} \hat{b}_{j,2} + ... +
\epsilon_j,
\quad
\epsilon \sim \mathrm{Normal}(0, \sigma^2),
\quad \mathrm{for}\ j  =  1,...,m.
\end{equation}

Our two machine learning superensemble models, a random forest and a
generalized boosted model (GBM), were based on regression trees. Regression
trees sequentially determine what value of a predictor best splits the
response data into two 'branches' based on a loss function [@breiman1984]. In
random forests, a series of regression trees are built on a random subset of
the data and random subset of the covariates of the model [@breiman2001]. In
GBMs, each subsequent model is fit to the residuals from the previous model;
data points that are fit poorly in a given model are given more weight in the
next model [@elith2008]. Random forests and GBMs can provide strong predictive
performance and fit highly non-linear relationships [@elith2008; @hastie2009].
We fit random forest models with the `randomForest` package [@liaw2002] for R
with the default argument values. We fit boosted regression tree models with
the `gbm` package [@ridgeway2015] for R. We fit GBMs with 2000 trees, an
interaction depth of $6$, a learning rate (shrinkage parameter) of $0.01$, and
all other arguments at their default values.

@@

Superensemble models allow us to incorporate additional covariates and
potentially leverage interactions between these covariates and individual model
predictions. Additional covariates could be, for example, life-history
characteristics, information on exploitation patterns, or statistical properties
of the data. For simplicity, and since many possible covariates, such as fishing
effort dynamics, would not be known for many real-world data-limited scenarios,
we added only one set of additional covariates: spectral properties of the catch
time series. Spectral analysis decomposes a time series into the frequency
domain and provides a means of describing the cyclical shape of the catch series
that is independent of time series length (except in affecting precision) and
independent of absolute magnitude of catch. We fit spectral models to the
scaled catch time series (catch divided by maximum catch) with the `spec.ar`
function in R and recorded representative short- and long-term spectral
densities at frequencies of 0.20 and 0.05, which correspond to 5- and 20-year
cycles. We include the results of adding these additional covariates in the
supplementary materials.

## Applying the superensemble models and testing performance

Once the superensemble models are built, we can apply them to new stocks that
we'd like to know the status of. To do this we applied the individual models
to our stocks of interest and then applied these data to our already built
superensemble models. In this paper we applied the superensemble models to
subsets of the simulated data as a cross validation test and to the RAM Legacy
Stock Assessment database to test on real stocks.

We used repeated three-fold cross validation to test predictive performance: we
randomly divided the dataset into three sets, built superensemble models on
two-thirds of the data, and evaluated predictive performance on the remaining
third. We repeated this across each of the three splits and then repeated the
whole procedure 50 times to account for bias that may result from any one set
of validation splits. In the simulated dataset, there were 10 replicates of
each unique combination of simulation parameters that differed only in
stochastic variability. Since the dynamics of these populations were often
similar, we grouped these stocks in the cross-validation process into either
the training or testing split.

We also tested our ensemble methods on the RAM Legacy Stock Assessment Database
--- a compilation of stock-assessment output from hundreds of exploited marine
populations around the world. Our analysis of the stock-assessment database was
based on version 2.5. After removing stocks for which at least one of the
individual models described below did not converge (121), this database
included `r ram_stocks_n` stocks. We removed these stocks for all methods ---
both for the individual and superensemble models. An alternative would be to
fit superensemble models to subsets of the individual models that did converge,
but for the purposes of this paper we will only use superensemble models that
use all four individual models.

In the case of the RAM Legacy Stock Assessment Database, we used superensembles
trained on the entire simulation dataset. However, since mPRM is built on the
same stock assessment database, we applied three-fold cross-validation to the
data underlying the mPRM model so that the dataset with which mPRM was trained
(for the individual model and superensemble) was separate from the dataset with
which it was tested.

Predictive performance can be evaluated with metrics that represent a variety
of modelling goals. For continuous response variables such as the mean and
slope of population status, performance metrics often measure some form of
bias, precision, accuracy (a combination of bias and precision), or the ability
to correctly rank or correlate across populations [e.g. @walther2005]. Here, we
measure proportional error, defined as $(\hat{\theta} - \theta)/\abs{\theta}$,
where $\hat{\theta}$ and $\theta$ represent estimated and 'true' (or
stock-assessed) mean or slope of $B/B_\mathrm{MSY}$. We calculated median
proportional error to measure bias, median absolute proportional error to
measure accuracy, and Spearman's rank-order correlation between predicted and
'true' values to measure the ability to correctly rank populations. When
testing with the RAM Legacy Stock Assessment database, we treated the
estimates from these data-rich stock assessments as known without error. Thus,
any error in the stock assessment estimates of the mean or slope of
$B/B_\mathrm{MSY}$ also contributes to our estimates of prediction error for
each of the four data-limited models and the ensembles. Code to reproduce our
analysis and an R package to fit the data-limited assessment models are
available at <https://github.com/datalimited/ensembles> (*Private while in
review*).

@@

<!--> TODO update the result numbers to report the models that don't have the covariates -->

@@

# Acknowledgments

We thank members of Phase I of the working group 'Developing new approaches to
global stock status assessment and fishery production potential of the seas'
who contributed to developing the data-limited methods and simulations used in
our analysis. We thank E. Jardim, F. Scott, and J.A. Hutchings for helpful
comments during the development of this project, and R.D. Methot for comments
on an earlier version of the manuscript. We thank the Gordon and Betty Moore
Foundation for funding the working group 'Applying data-limited stock status
models and developing management guidance for unassessed fish stocks'.

@@

<!-- vim: set formatoptions=nroq: -->
